Rising from the Ashes II: The Bar-driven Abundance Bimodality of the Milky Way

### 2025-02-19 paper/ms.tex
Fix Brook et al citations and add Lu/Ratcliffe citations

The origin of the bimodality is a topic of active debate, with three main scenarios proposed to explain it. First, it is a result of internal secular processes that generate the bimodality through radial migration \citep{2009MNRAS.396..203S,2021MNRAS.507.5882S,2023MNRAS.523.3791C} or clump formation \citep{2019MNRAS.484.3476C,2020MNRAS.492.4716B,2021MNRAS.502..260B,2023ApJ...953..128G}. Second, the bimodality is generated through gas infall scenarios, either from specific gas accretion episodes from the intergalactic medium \citep{1997ApJ...477..765C,2009IAUS..254..191C,2017MNRAS.472.3637G,2019A&A...623A..60S}, or through a more self-consistent collapse sequence of the circumgalactic medium driven through feedback \citep{2021MNRAS.501.5176K}. \edit1{Third and finally, the bimodality is generated through a merger process, either by enhancing the star formation rate (SFR) of the Galaxy \citep{2018MNRAS.474.3629G}}\footnote{\edit1{\citet{2004ApJ...612..894B,2005ApJ...630..298B,2007ApJ...658...60B,2010MNRAS.402.1489R} also explored the $\alpha$-enhancement of the thick disk resulting from gas-rich mergers. However, they did not explore the arising of a clean separation purely in chemistry.}} or by supplying a relatively pristine gas supply that resets the metallicity of the Galaxy \citep{2020MNRAS.491.5435B,2024MNRAS.528L.122C}. Strong evidence that the Milky Way did undergo a merger with the Gaia-Sausage-Enceladus (GSE) satellite supports the merger-related scenarios \citep{2018MNRAS.478..611B,2018Natur.563...85H,2020ApJ...901...48N,2024ApJ...972..112C}.

>>>>

The origin of the bimodality is a topic of active debate, with three main scenarios proposed to explain it. First, it is a result of internal secular processes that generate the bimodality through radial migration \citep{2009MNRAS.396..203S,2021MNRAS.507.5882S,2023MNRAS.523.3791C} or clump formation \citep{2019MNRAS.484.3476C,2020MNRAS.492.4716B,2021MNRAS.502..260B,2023ApJ...953..128G}. Second, the bimodality is generated through gas infall scenarios, either from specific gas accretion episodes from the intergalactic medium \citep{1997ApJ...477..765C,2009IAUS..254..191C,2017MNRAS.472.3637G,2019A&A...623A..60S}, or through a more self-consistent collapse sequence of the circumgalactic medium driven through feedback \citep{2021MNRAS.501.5176K}. \edit1{Third and finally, the bimodality is generated through a merger process, either by enhancing the star formation rate (SFR) of the Galaxy \citep{2018MNRAS.474.3629G}}\footnote{\edit1{\citet{2004ApJ...612..894B,2005ApJ...630..298B,2007ApJ...658...60B,2010MNRAS.402.1489R} also explored the $\alpha$-enhancement of the thick disk resulting from gas-rich mergers. However, they did not explore the arising of a clean separation purely in chemistry.}} or by supplying a relatively pristine gas supply that resets the metallicity of the Galaxy \citep{2020MNRAS.491.5435B,2024MNRAS.528L.122C}.

\edit1{The merger hypothesis is supported by strong evidence that the Milky Way did undergo a merger with the Gaia-Sausage-Enceladus (GSE) satellite supports the merger-related scenarios \citep{2018MNRAS.478..611B,2018Natur.563...85H,2020ApJ...901...48N,2024ApJ...972..112C}. In deriving stellar birth radii from assuming a linear relation between \FeH{} and radius with a time-evolving slope, \citet{2024MNRAS.535..392L} argued that the bimodality resulted from a steepening of the metallicity gradient at the time of the GSE merger \citep[see also][]{2023MNRAS.525.2208R}. This provides further evidence that the bimodality formed around the same time as the GSE merger, although the case for a causal connection is less clear.}
